### 2016-01-06 papers/CAFe/HIPS-paper/coarray-comparison.tex

We emphasize some of these advantages by comparing the \texttt{Laplace} implementation with the
implementation of the same algorithm from the original Numrich and Reid
paper\cite{Numrich:1998:CFP:289918.289920} first describing coarrays in Fortran.  We should point
out that this comparison is somewhat unfair, because Numrich and Reid were introducing coarray
notation for transferring memory on distributed memory architectures, not demonstrating how ideally
one should use coarrays within a large application.

@@

There are several advantages to LOPe, including:

@@

\item
LOPe requires the implementation of the algorithm to be separate from the call to effect the halo
transfer.  Removing boundary condition specification (e.g., the cyclic boundary conditions
implemented in the CAF example) from the algorithm allows the boundary conditions to be changed
without changing algorithm code.

\item
LOPe applies the transfer of halo memory across possibly multiple levels of memory with the LOPe
intrinsic \texttt{TRANSFER\_HALO} function.  Thus the LOPe algorithm can be run on a machine with
many interconnected nodes, each containing hybrid processor cores.  The coarray example can only be
run on multiple nodes without accelerator cores.

@@

\item
The algorithm implementation is separate from any specification as to where the algorithm
is to be executed.  The CAF example explicitly denotes where a statement is to executed
with a control flow construct like \texttt{if (me == 1)}.

@@

\item
The semantics of LOPe makes explicit management of array temporaries (e.g., \texttt{u} and
\texttt{new\_u} by the programmer unnecessary (though still possible).  Because in LOPe the
halo region is a language construct, the compiler is better able to manage temporary
buffers than users on the target hardware platform.

@@

\item
A programmer can't make synchronization errors in calls to LOPe functions as synchronization is
implicit in the \texttt{CONCURRENT} attribute.  A thread running a concurrent procedure is provided
with a copy of it's local array element plus halo that is consistent with the state of memory
\emph{at the time of invocation of the procedure}.  Stores to an individual thread's local array
element (by that thread) are never visible to other threads.  Normal coarray programs (like MPI)
require explicit synchronization to ensure that processes arrive at the same program location
before a memory read occurs (for example).  LOPE encourages the creation of small functions and
lets the compiler fuse the functions together to improve performance and to provide necessary
synchronization.

>>>>

We emphasize some of these advantages by comparing the \texttt{Laplacian} implementation in Fig. 1 with the
implementation of the same algorithm from the original Numrich and Reid
paper\cite{Numrich:1998:CFP:289918.289920} first describing coarrays in Fortran.  We should point
out that this comparison is somewhat unfair, because Numrich and Reid were introducing coarray
notation for transferring memory on distributed memory architectures, not demonstrating how ideally
one should use coarrays within a large application.

@@

%%There are several advantages to LOPe over the CAF example, including:
A comparison of Fig. 1 to the CAF example suggests the following advantages:

@@

\item
LOPe requires the implementation of the algorithm to be separate from the call to effect the halo
transfer.  Removing boundary condition specification
%%(e.g., the cyclic boundary conditions implemented in the CAF example)
from the algorithm allows the boundary conditions to be changed
without changing algorithm code.

\item
LOPe applies the transfer of halo memory across possibly multiple levels of memory with the LOPe
intrinsic \texttt{TRANSFER\_HALO} function.  Thus the LOPe algorithm can be run on a machine with
many interconnected nodes, each containing hybrid processor cores.
%%The coarray example can only be run on multiple nodes without accelerator cores.

@@

\item
The algorithm implementation is separate from any specification as to where the algorithm
is to be executed.  The CAF example explicitly denotes where a statement is to be executed
with the control flow construct \texttt{if (me == 1)}.

@@

\item
The semantics of LOPe makes explicit management of array temporaries (e.g., \texttt{u} and
\texttt{new\_u}) by the programmer unnecessary.
%%(though still possible).
Because in LOPe the
halo region is a language construct, the compiler is better able to manage temporary
buffers than users on the target hardware platform.

@@

\item
A programmer can't make synchronization errors in calls to LOPe functions as synchronization is
implicit in the \texttt{CONCURRENT} attribute.  A thread running a concurrent procedure is provided
with a copy of its local array element plus halo that is consistent with the state of memory
\emph{at the time of invocation of the procedure}.  Stores to an individual thread's local array
element (by that thread) are never visible to other threads.  Normal coarray programs (like MPI)
require explicit synchronization to ensure that processes arrive at the same program location
before a memory read occurs (for example).  LOPE encourages the creation of small functions and
lets the compiler fuse the functions together to improve performance and to provide necessary
synchronization.
